Impact of switching bug trackers: a case study on a medium-sized open source project

### 2019-03-05 article/article.tex
Article update.

\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
\usepackage{amsmath}
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage{hyperref,xcolor}
\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Impact of switching bug trackers: a case study}
%
%\titlerunning{}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Th\'eo Zimmermann\inst{1,2} \and
Annal\'i Casanueva Art\'is\inst{3}\thanks{This work was partly funded by El Conacyt of the Mexican government.}}
%
\authorrunning{T. Zimmermann and A. Casanueva Art\'is}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{$\Pi R^2$, Inria, Paris, France \and
IRIF, Paris-Diderot University, Paris, France\\
\email{theo@irif.fr} \and
Paris School of Economics, Paris, France}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Any significant software project requires a bug tracker. In open source development, this plays an even bigger role. The impact of the bug reporting environment on the bug tracking activity is difficult to assess. In this paper, we take advantage of the switch from Bugzilla to GitHub of the bug tracker of Coq, a medium-sized open source project, to evaluate the impact that such a change can have. 
We first report on the switch itself, in particular the migration of 4900 preexisting bug reports using the GitHub issue import API; then we analyze data from before and after the migration. We show that the switch induces 
an increase in bug reporting, particularly from principal developers themselves, and more generally an increased engagement with the bug tracking platform, with more comments by developers and also more external commentators.

@@

\keywords{bug tracker \and switch \and migration  \and bug report \and issue \and GitHub \and Bugzilla \and open source.}
\end{abstract}
%
%
%
\section{Introduction}

Bug reporting is an essential part of software development. However, it involves
a non-negligible human component, particularly in the context of an open source project where a larger part of bugs are reported by independent users and therefore depend a lot on the good will of individuals. Creating an environment that makes bug reporting easier and more appealing may lead to an increase of bug reporting (from both developers and independent users) which should ultimately benefit  software development \cite{Bettenburg2008} (provided that developers are equipped to handle the incoming reports \cite{Anvika,Kanwal2012}). However, the impact of such change has rarely been studied. Indeed, comparing the number of bug reports in different environments is generally not compelling because it is likely to depend  not just on the reporting environment but also on each particular software and its development dynamics. To do a valid comparison one would ideally need to compare two bug reporting environments for the same software, in the same period. But, or course, we can't observe this ideal. A context approaching this ideal scenario would be to compare the number of bug reports of a particular software before and after the switch to a new environment. Yet, substantial changes in the environment of bug reporting (e.g. changes of bug tracker platform) are rare because once a bug system is in place, any change comes with a significant cost.

We use the switch, from Bugzilla to GitHub, of the bug tracker system of the Coq proof assistant, a medium-sized open source software, to analyze the  effects of a change in the bug reporting environment on the number of bugs, reporters, bug comments and commentators. We describe the switch event and the associated migration process for preexisting bugs. Thanks to the way this migration was conducted, getting data on bug reporting from before and after the switch is easily done using the GitHub GraphQL API. Using a Regression in Discontinuity analysis we find strong evidence that the switch of platform increased the number of bug reports; started an increase of bug reporters per period (i.e. in a certain period there are more distinct reporters); increased the number of comments by developers and the number of distinct commentators per period among non-developers. Taken together our results suggest that switching from Bugzilla to GitHub increased the number of bug reports and bug reporters (probably by making the bug reporting process easier) but also increased the openness of software development and aroused discussion around bugs (notably between developers and non-developers).

@@

\subsection{Coq and GitHub}
Coq is a (free and open source) proof assistant, that is a software to write and automatically verify mathematical proofs (and programs). It originated as a research project in 1984 and the development team is still composed, for the most part, of researchers. It gained in popularity along the years, and is now at the center of many verification projects (some very large academic projects, and some industrial projects). Its initial developers were awarded  the ACM SIGPLAN Programming Languages Software 2013 Award and the ACM Software System 2013 award.

The development is currently centered around the GitHub platform. The current repository of Coq the current repository counts more than 25000 commits. Some of these commits date back from 1999, when the code received major architectural changes. The GitHub repository was initially created as a mirror of the official repository (which itself was migrated from SVN to git in 2013) and really started attracting the interest of the development team as a way to open up the development to external contributions, especially starting with the first Coq coding sprint in 2015 (later renamed into the Coq Implementors Workshop and held every year in France). From then on, core developers started using pull requests for their own changes more and more often, and it became the norm around the beginning of 2017 when systematic continuous integration tests were introduced on every pull requests. In the last two years, more than 2,200 pull requests have been opened.

@@

From 2007 to 2017, the Coq bug tracker platform was a self-hosted Bugzilla instance (before this, from 2001 to 2007 it was using JitterBug, a now discontinued bug tracking system, and before 2001 a simple mailing list). In 2017, developers had become quite accustomed to the look and feel of discussion threads on GitHub pull requests, which offered a much nicer experience than the Bugzilla bug tracker. Younger developers (in particular, one of the author of this article, who eventually pushed for the switch and conducted the migration) were frustrated by what they perceived as an older and heavier platform. The rest of the developers were convinced to migrate towards a non-self-hosted solution in August 2017, when the whole Coq website was shut down for more than a week  because an IT team was frightened by a scam bug report.\footnote{\url{https://sympa.inria.fr/sympa/arc/coq-club/2017-08/msg00040.html}}

The expected advantages of the GitHub issue tracker over Bugzilla were:
\begin{itemize}
\item Better formatting, ability to edit comments, and more generally a more pleasant tool to use. We felt that Bugzilla was unpleasant enough to discourage developers to use it to its full potential.
\item Browsing between code, pull requests and bug reports is easier when everything is in the same place. Furthermore, it means a common notification, mention, and assignment system.
\item Better indexing by search engines.
\item Shared management of permissions, milestones, categories (labels), etc.
\item Easier to get started for newcomers (especially as many may already know GitHub and already have a GitHub account).
\end{itemize}

The possible drawbacks were mostly that GitHub provides a less-advanced system with very few customization options. In particular:
\begin{itemize}
\item The permission system is much less fine-grained. Giving triaging rights to users requires giving them ``write access'', which includes commit rights on non-protected branches. That's why the Coq GitHub repository now only has protected branches (and accidentally-pushed topic branches are promptly deleted).
\item Creating a new bug report doesn't require filling a form, but typing a message. Fortunately, issue templates help give some hints towards the expected structure of the report, and being able to strip this template can actually give flexibility to developers who might open new issues for concerns that do not qualify as bugs.
\item There is no import mechanism that allows to keep the number and author of preexisting reports. Fortunately, as we will see in Sec.~\ref{migration}, it was still possible to import bug reports while keeping most of the meta-information and renumbering the fewest possible bug reports. This was an important requirement for the switch.
\end{itemize}

@@

Given the advantages of the GitHub issue tracker that were listed above, we expected the switch to:
\begin{itemize}
\item increase the number of bug reports, by developers in particular;
\item increase the diversity of bug reporters;
\item improve the interaction between reporters and developers.
\end{itemize}

@@

The only complicated part of the bug tracker switch was the migration of preexisting bug reports. We reused a tool by Andriy Berestovskyy\footnote{Tool available at \url{https://github.com/berestovskyy/bugzilla2github}.} which is designed to import Bugzilla reports (extracted as an XML dump) to GitHub using its REST API. The bugs are imported in an order designed to preserve numbers whenever possible. Bugs whose number is unavailable (e.g. because the number is already taken by a GitHub pull request) are postponed and renumbered. We implemented several improvements to make the tool better fit our needs:
\begin{itemize}
\item \textit{Allowing non-consecutive bug numbers:} Our imported set of bug reports had some holes in the numbering due to deleted bugs. We use postponed bugs to fill the holes. This improvement has now been integrated upstream.
\item \textit{Saving a table of correspondence for renumbered bugs:} This was used later to redirect the old bug URLs to the new ones.
\item \textit{Using the GitHub issue import beta API and overcoming the GitHub rate limits:} Creating a new issue or a new comment through the normal GitHub REST API will trigger notifications (for people who are watching the repository or are mentioned in the issue thread). Such actions are therefore subject to a strict rate limit which prevented using this tool for importing more than a few hundred bugs. Fortunately, GitHub provides a (beta) issue import API which, in addition to not triggering notifications, also allows importing one bug, its comments, and meta information such as closed status and assignee in a single request (thus reducing the duration to import 4900 bugs to just a few hours). Finally, using this API allows to keep the dates of imported bug reports and comments, which is very useful to this study. We didn't manage to import correct closing dates, so we won't study the impact of the bug tracker switch on the time it takes to close a bug.
\end{itemize}

This tool required a correspondence table between Bugzilla and GitHub accounts. Ours was created by manually matching 217 (out of 686) Bugzilla accounts to 175 GitHub accounts,\footnote{The correspondence table can be seen at \url{https://gist.github.com/Zimmi48/d923e52f64fe17c72852d9c148bfcdc6##file-bugzilla2github-L47-L264}.} the difference being due to the high prevalence of duplicate Bugzilla accounts. We gave priority to finding GitHub accounts for people who still had opened bug reports. The most efficient way to do the matching was by searching the name of the reporter + ``github'' on Google. We also sent e-mails to people we hadn't managed to match (and who still had opened bugs).

\paragraph{Urgency.}
Once the switch was approved by the development team on October 3-4\textsuperscript{th}, 2017, the migration had to happen as soon as possible because every new pull request before the migration added to the number of bug reports that would need to be renumbered. It was conducted\footnote{More details on the process are available in a blog post at \url{http://theoz.im/bugzilla}. } on October 18\textsuperscript{th},  the day after the 8.7.0 release (and not before to avoid disturbing the release process). Only 502 bugs (whose numbers were below 1154) had to be renumbered. This number is quite low because these bugs are dating back from the JitterBug era and during the previous switch only the bugs that were still open had been migrated.

\section{Data}

@@

All the data for this study was extracted on September 14\textsuperscript{th}, 2018 using the GitHub GraphQL API. Using this API allows to do large requests (100 nodes in a single request) with just the information we need (thus both reducing the bandwidth usage and speeding up the extraction process). We provide the extracted data as CSV files and a Jupyter notebook\footnote{Available at \url{https://github.com/Zimmi48/impact-of-switching-bug-trackers}.} with the code to request this data from GitHub, to load the CSV files, and to run the analyses.

Migrated bugs and comments had their author information stored in the text so we extract it from there. Dates of bug and comment creation have been preserved so we can obtain them transparently for bugs before and after the migration. Comments dating from the JitterBug era do not have valid author information (and closed bug reports from this era have not been migrated) so we won't consider any data from before 2008.

@@

\paragraph{Excluding specific reporters.}
We have one specific reporter, Jason Gross, who is alone responsible for almost a quarter of all bug reports. To avoid having the behavior of a single individual strongly impact the overall statistics, we exclude his comments, bug reports and pull requests and all the comments they received from our analysis. Similarly, we also exclude the developer who was the initiator and conductor of the bug tracker switch and migration and who is an author of this article.

\paragraph{Merging duplicate Bugzilla accounts.}
This is only relevant to the new reporter analysis which itself is only detailed in the accompanying Jupyter notebook, so refer there for more information.

\paragraph{Removal of migration artifact comments.}
The migration tool created an issue whose body contains only meta-information, and a first comment with the description, for every bug report. These comments are thus migration artifacts and we remove them from our comment analyses. They can be easily identified because they are the comments that were posted at the exact same date and time as the corresponding bug report.

@@

Our outcome variables will be the numbers of bug reports, distinct bug reporters, new bug reporters (who had never reported a bug before), comments and distinct commentators, for a given period. We will use 15-day periods to have sufficiently many bugs by period. Indeed, in our data we have an average of 13 bug reports by 15-day period; 50\% of the periods have between 7 and 18 reports; 90\% of the periods have between 4 and 27. This also allows to have 40 observations centered around the date of the switch for our Regression in Discontinuity analyses.

We will analyze heterogeneous effects by distinguishing between developers and others (others being assimilated to ``users''). We define ``developers'' as the persons who have contributed more than 100 commits since 2008.

When studying the influence of new releases, we define the release dates by the dates of the corresponding news items on the official website (\url{https://coq.inria.fr/news}). We selected beta and major releases and excluded release candidates and patch-level releases. The first release to appear in the news section of the current official website is the 8.2beta release in June, 2008.

\section{Descriptive statistics}

The accompanying Jupyter notebook contains lots of descriptive statistics and figures, such as the distribution of bug reporting in the year, and in the day, graphs of bug reporting for individual developers, etc. We only present a subset of these figures.

@@

Fig.~\ref{cumulative_bugs_and_comments} gives a first (cumulative) view of two of our outcome variables: the number of bug reports, and the number of comments. From this figure, we can clearly notice the dominance of developers in the number of bug comments and the dominance of non-developers in the number of bug reports. We can also already notice some slope changes at the bug tracker switch date. The magnitude and significance of this changes will be assessed in the remainder of this article.

>>>>

%\documentclass[runningheads]{llncs}
\documentclass[conference]{IEEEtran}
%\IEEEoverridecommandlockouts
%
\usepackage{graphicx}
% If possible, figure files should be included in EPS format.
%
\usepackage{amsmath}
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\usepackage{hyperref,xcolor}
\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Impact of switching bug trackers: a case study on a medium-sized open source project}
%
\author{Th\'eo Zimmermann\\
$\Pi R^2$, Inria and
IRIF, Paris-Diderot University\\
Paris, France\\
Email: theo@irif.fr
\and
Annal\'i Casanueva Art\'is\thanks{This work was partly funded by El Conacyt of the Mexican government.}\\
Paris School of Economics\\
Paris, France}
%
\maketitle
%
\begin{abstract}
For most software projects, the bug tracker is an essential tool. In open source development, this tool plays an even more central role as it is generally open to all users, who are encouraged to test the software and report bugs. Previous studies have highlighted the act of reporting a bug as a first step leading a user to become an active contributor.

The impact of the bug reporting environment on the bug tracking activity is difficult to assess because of the lack of comparison points. In this paper, we take advantage of the switch, from Bugzilla to GitHub, of the bug tracker of Coq, a medium-sized open source project, to evaluate the impact that such a change can have.

We first report on the switch itself, in particular the migration of 4900 preexisting bug reports using the GitHub issue import API. Then we analyze data from before and after the switch using a Regression on Discontinuity analysis, a novel methodology in the context of empirical software engineering.
We show that the switch induces an increase in bug reporting, particularly from principal developers themselves, and more generally an increased engagement with the bug tracking platform, with more comments by developers and also more external commentators.
\end{abstract}

@@

\begin{IEEEkeywords}
bug tracker, switch, migration, bug report, issue, GitHub, Bugzilla, open source, data mining, Regression on Discontinuity
\end{IEEEkeywords}
%
%
%
\section{Introduction}

Bug reporting is an essential part of software development.
In the context of an open source project, the bug reporting and fixing process is generally done on a public bug tracking platform, and, in the absence of paid testers, it depends a lot on the good will of independent users. Therefore, it involves a strong social component.

However, having an open bug tracking system is not enough to get the bugs reported. In addition to the software having enough users, the process of bug reporting must be easy and appealing. Therefore, creating an environment that eases bug reporting and makes it more appealing may lead to an increase in bug reporting activity (from both developers and independent users), which is well associated with an increase in software quality. First, if the software has not changed, receiving more bug reports indicates that more defects have been found, and not a sudden apparition of defects (Linus's Law ``given enough eyeballs, all bugs are shallow''~\cite{raymond1999cathedral,van2009shallow}). Moreover, assuming that developers are equipped to cope with the increased number of incoming reports~\cite{anvik2005coping, davidson2011coping}, even duplicate reports are not necessarily harmful~\cite{Bettenburg2008}. Second, more activity on the bug tracker can also mean users are helping to reproduce bugs, produce traces, etc. and thus are working with the developers to get the bugs fixed~\cite{breu2010information}. More generally, reporting new bugs and discussing existing bug reports has been shown to be an important step on the path to becoming an active contributor of an open source project~\cite{jensen2007role, nakakoji2002evolution, ye2003toward}.

Despite the importance of the bug tracking environment that we have just highlighted, the impact of a change in this environment has rarely been studied. It is generally difficult to compare the number of bug reports in different environments because it is likely to depend  not just on the reporting environment but also on each particular software and its development dynamics, which would make such an analysis not compelling. To do a valid comparison, one would ideally need to compare two bug reporting environments for the same software, used by the same users, to report the same bugs, in the same period. This is of course impossible. A context approaching this ideal scenario would be to compare the number of bug reports of a particular software before and after the switch to a new environment. Yet, substantial changes in the environment of bug reporting (e.g. changes of bug tracker platform) are rare because once a bug tracking platform is in place, any change comes with a significant cost.

In this paper, we study the case of the Coq proof assistant, a medium-sized open source software project. We use the switch, from Bugzilla to GitHub, of the project's bug tracker, to analyze how a change in the bug reporting environment affects the number of bugs, reporters, bug comments and commentators. Using a Regression on Discontinuity design (RDD)~\cite{angrist2008mostly,angrist2014mastering,jacob2012practical,lee2010regression}, we find strong evidence that the change of platform increased the number of bug reports and comments by developers, and increased the number of distinct commentators each week among non-developers. Our results suggest that switching from Bugzilla to GitHub made the bug reporting process easier and more appealing but also opened the software development to more external contributors and stimulated discussion around bugs (notably between developers and non-developers).

The contributions of this paper are practical, empirical, and methodological. First, we improved an existing bug tracker migration tool to handle thousand of bug reports while preserving meta-data, and we helped the Coq development team migrate their bug tracker from Bugzilla to GitHub. Second, we analyze the causal impact of this switch on the bug tracking activity. Third, we introduce to the field of software engineering a state-of-the-art method for quantitative public policy evaluation: Regression in Discontinuity Design. In particular, we explain this method to give the basic intuition behind; we point the reader to the both introductory and state-of-the-art literature and we illustrate one application.

The rest of this paper is organized as follows: In Sec.~\ref{related-work}, we discuss related work. In Sec.~\ref{context}, we present the Coq project and the context of the switch. In Sec.~\ref{hypotheses}, we formulate hypotheses on the impact of the switch. In Sec.~\ref{migration}, we explain the migration process. In Sec.~\ref{data}, we describe the data collection and pre-processing method, and we list the variables of interest. In Sec.~\ref{descriptive-stats}, we give a preliminary view of the bug tracking activity. In Sec.~\ref{methodology}, we explain the Regression on Discontinuity methodology, of which we present the results in Sec.~\ref{results}. We interpret the results and relate them to our hypotheses in Sec.~\ref{discussion}. We discuss the threats to validity and  present robustness checks in Sec.~\ref{threats}. We conclude in Sec.~\ref{conclusion}.

\section{Related work}

\label{related-work}

While there is a very large literature on many aspects of bug reporting (see~\cite{strate2013literature, zhang2016literature} for literature reviews), there is a lack of literature measuring the influence of the bug reporting environment on the bug reporting activity or on any other aspects of software development. To our knowledge, there is no previous work measuring the impact of a change in bug reporting environment. More generally, there is little literature that compare bug trackers.

\subsection{Comparing and proposing features of bug trackers}

A simple, but quite limited, way to compare bug trackers is to look at the features that each of them proposes.

Karre et al.~\cite{karre2017does} compared 31 bug trackers feature-wise, and gathered these tools in four clusters. Bugzilla belongs to the cluster of bug trackers with many features (together with RedMine and Mantis), while GitHub belongs to a cluster of bug trackers attached to source code management systems (together with Savannah and BitBucket).
%Bugzilla has 20 of the 23 features they listed, and scores the highest number of features (ex. aeqo with RedMine). They are closely followed by Mantis, and two other bug trackers, with 19 features. GitHub only has 15 features out of the 23 listed (overall 11 listed systems have more features).

In the case of the Coq project, the loss of some specific features from Bugzilla was one of the arguments against the switch. Nonetheless, GitHub has some features that Bugzilla hasn't, and it is clearly not the number of features that determines which software is better or more adapted to the specific use case.

%We will note that GitHub has very recently added a number of new features to the issue tracker, but these were not present at the time of the switch (just-in-time duplicate retrieval, issue deletion, transfer and pinning, support for multiple issue templates, etc.).

Additionally, Karre et al. recommended some missing features, including better support for migrating bug tracking data from a system to another. Our work has contributed to improve the situation in that sense. Nonetheless, it would be preferable if there existed a standard structured format for storing bug tracking information and all major bug tracking systems supported exporting to and importing from this format.

Abaee and Guru~\cite{abaee2010enhancement} listed the features of four commercial bug trackers, and proposed their own bug tracker with some new features, but without any evaluation.

There is much more literature proposing new features to add to bug trackers, but they are rarely evaluating the impact of adding such features on the bug tracking activity, even when a prototype was presented.

Baysal et al.~\cite{baysal2013situational} identified the need for personalized issue tracking systems after interviewing twenty Mozilla developers. They proposed a Bugzilla extension to address this need, and gave a, mostly qualitative, assessment by interviewing developers using their tool in~\cite{baysal2014no}.

Bortis~\cite{bortis2016porchlight} developed a bug triaging tool and evaluated how users interacted with it. However, there was no evaluation of its impact in the context of an actual software project.

Just et al.~\cite{just2008towards} surveyed developers from three large open source projects (Apache, Eclipse, and Mozilla; all three projects use Bugzilla) to identify missing features and provided recommendation to design better bug tracking systems.
%They remark that bug trackers are often mere interfaces to bug databases and ask too much information from end-users without guiding them. They recommend providing guidance to bug reporters depending on their level of expertise. This issue is indeed something that we have noticed with Bugzilla and other bug trackers of the same category. Users are encouraged to self-triage and often do so inaccurately. On the contrary, GitHub only provides title and body fields to bug reporters and leaves triaging (labeling) to the contributors with write access. Nonetheless, the issue templates can be used to fill the bug report body field with advice on how to present a bug report. GitHub has recently added support for multiple issue templates which would allow to let the reporter choose between, for instance "My first bug report", "My first feature request", "Bug report (experienced reporter)", and "Feature request (experienced reporter)", with varied requirements and levels of explanations...

Our article goes beyond a descriptive or normative perspective and quantitatively measures the causal impact of the change of the bug tracking platform (a crucial part of the bug tracking environment) on various aspects of bug reporting activity.

\subsection{Analysis of projects' bug tracking data}

Another way of comparing bug tracking systems could have been to conduct large-scale studies of many software projects using various bug trackers and provide some clues of the differences induced by the use of the different bug tracking systems. However, there are no such comparative studies in the literature.

Sowe et al.~\cite{sowe2013multi} noted that most preceding literature had only been comparing few projects at a time, usually using a single bug tracking system. In their study, they addressed this in part by studying hundreds of projects, but they did not mention which bug trackers the projects they compared used.\looseness=-1

Bissyandé et al.~\cite{bissyande2013got} published the same year a larger scale investigation using ten thousand projects' bug tracking data, analyzing such things as the correlation between a project's success and its bug tracking activity. All of the projects studied were using the same bug tracker (GitHub's).

Francalanci and Merlo~\cite{francalanci2008empirical} analyzed the bug fixing process by studying closed bugs from nine open source projects (four using JIRA, four using SourceForge's bug tracker, and one using another bug tracker). They did not, however, try to correlate the bug tracking activity with the bug tracking system used.

%According to the survey in~\cite{karre2017does}, Bugzilla is the most used bug tracker (respondent were 32\% to use Bugzilla and 15\% to use GitHub). There is an obvious selection bias due to the choice of open source project contributors this survey was sent to. Overall, given the sheer number of projects hosted on GitHub, it does not make any doubt that the GitHub issue tracker is the most used bug tracker in number of projects. But this raw fact does not matter much, given that very small projects will use their bug tracker very little. 
\subsection{Switching developer tools}

While, to the best of our knowledge, there is no work evaluating the impact of switching bug trackers, some studies have been conducted to evaluate the impact of switching other developer tools. Since bug trackers are important developer tools, our article also relates and contributes to this literature.

De Alwis and Sillito~\cite{de2009software} surveyed developers to understand the perceived impact of moving from centralized to decentralized version control systems. As with bug trackers, the move is decided because of anticipated benefits: among them is the openness to external contributors. But the switch incurs challenges in the migration process (in particular, developers are very concerned to preserve the full history of the project) and some features are lost in the new tool (in particular, the monotonically increasing revision numbers).
Muşlu et al.~\cite{mucslu2014transition} performed a similar study based on interviews and a survey. They identified expectations and barriers and provided recommendations to developers wanting to conduct such a migration.

Squire~\cite{squire2015should} measured the impact of switching developer support channels from mailing lists and self-hosted forums to Stack Overflow, by first identifying the expectations of 20 projects which had officially decided on such a move, then by studying the impact on developer participation and response time for a selection of four such projects. On the other hand, Vasilescu et al.~\cite{vasilescu2014social} compared the behavior of the same users on the R mailing list and the Stack Overflow platform. Contrarily to a bug tracker which is expected to be centralized for a given project, support channels can be diverse. Project leaders can decide to abandon their mailing list to focus exclusively on Stack Overflow, but they do not decide to ``open'' a support channel on Stack Overflow: it can exist nevertheless and the switch can be organic.

@@

\label{context}

\subsection{Coq and GitHub}
Coq is a (free and open source) proof assistant, i.e., a software to write and automatically verify mathematical proofs (and programs). It originated as a research project in 1984 and the development team is still composed, for the most part, of researchers. It gained in popularity along the years, and is now at the center of many verification projects (some very large academic projects, and some industrial projects). Its initial developers were awarded  the ACM SIGPLAN Programming Languages Software 2013 award and the ACM Software System 2013 award.

% XXXX justify more the medium-size qualification

The development is currently centered around the GitHub platform. The Coq repository counts more than 25,000 commits. Some of these commits date back from 1999, when the code received major architectural changes. The GitHub repository was initially created as a mirror of the official repository (which itself was migrated from SVN to git in 2013) and really started attracting the interest of the development team as a way to open up the development to external contributions, especially starting with the first Coq coding sprint in 2015 (later renamed into the Coq Implementors Workshop and held every year in France). From then on, core developers started using pull requests for their own changes more and more often, and it became the norm around the beginning of 2017 when systematic continuous integration tests were introduced on every pull request. In the last two years, more than 2,200 pull requests have been opened.

@@

From 2007 to 2017, the Coq bug tracker platform was a self-hosted Bugzilla instance (before this, from 2001 to 2007 it was using JitterBug, a now discontinued bug tracking system, and before 2001 a simple mailing list). In 2017, developers had become quite accustomed to the look and feel of discussion threads on GitHub pull requests, which offered an arguably nicer experience than the Bugzilla bug tracker. Younger developers were frustrated by what they perceived as an older and heavier platform. The rest of the developers were convinced to migrate towards a non-self-hosted solution in August 2017, when the whole Coq website was shut down for more than a week  because an IT team was frightened by a scam bug report.\footnote{\url{https://sympa.inria.fr/sympa/arc/coq-club/2017-08/msg00040.html}}

The expected advantages of the GitHub issue tracker over Bugzilla were:
\begin{itemize}
\item Better formatting, ability to edit comments, and more generally a more pleasant tool to use. According to some Coq developers, Bugzilla was unpleasant enough to discourage developers to use it to its full potential.
\item Version control integration: Browsing between code, pull requests and bug reports is easier when everything is in the same place. Furthermore, it means a common notification, mention, and assignment system.
\item Better indexing by search engines.
\item Shared management of permissions, milestones, categories (labels), etc.
\item Easier to get started for newcomers (especially as many may already know GitHub and already have a GitHub account).
\end{itemize}

The possible drawbacks were mostly that GitHub provides a less advanced system with very few customization options. In particular:
\begin{itemize}
\item The permission system is much less fine-grained. Giving triaging rights to users requires giving them ``write access'', which includes commit rights on non-protected branches. That's why the Coq GitHub repository now only has protected branches (and accidentally-pushed topic branches are promptly deleted).
\item Creating a new bug report does not require filling a form, but typing a message. Fortunately, issue templates help give some hints towards the expected structure of the report, and being able to strip this template can actually give flexibility to developers who might open new issues for concerns that do not qualify as bugs.
\item There is no import mechanism that allows to keep the number and author of preexisting reports. Fortunately, as we will see in Sec.~\ref{migration}, it was still possible to import bug reports while keeping most of the meta-information and renumbering the fewest possible bug reports. This was an important requirement for the switch.
\end{itemize}

@@

\label{hypotheses}
%ambre: I think this section should be expanded a bit to add motivation for the hypotheses

Given the advantages of the GitHub issue tracker that were listed above, one could expect the switch to:
\begin{description}
\item[H1] Increase the bug tracking activity, in terms of the number of reported bugs;
\item[H2] Extend the pool of bug reporters and commentators;
\item[H3] Increase the bug tracking activity, in terms of the amount of interaction between users and developers.
\end{description}

@@

The only complicated part of the bug tracker switch was the migration of preexisting bug reports. We reused a tool by Andriy Berestovskyy\footnote{Tool available at \url{https://github.com/berestovskyy/bugzilla2github}.} which is designed to import Bugzilla reports (extracted as an XML dump) to GitHub using its REST API. The bugs are imported in an order designed to preserve numbers whenever possible. Bugs whose number is unavailable (e.g. because the number is already taken by a GitHub pull request) are postponed and renumbered. We implemented several improvements to make the tool better fit the needs of the Coq project:
\begin{itemize}
\item \textit{Allowing non-consecutive bug numbers:} The imported set of bug reports had some holes in the numbering due to deleted bugs. We use postponed bugs to fill the holes. This improvement has now been integrated upstream.
\item \textit{Saving a table of correspondence for renumbered bugs:} This was used later to redirect the old bug URLs to the new ones.
\item \textit{Using the GitHub issue import beta API and overcoming the GitHub rate limits:} Creating a new issue or a new comment through the normal GitHub REST API will trigger notifications (for people who are watching the repository or are mentioned in the issue thread). Therefore, GitHub chooses to impose a strict rate limit on these actions, which prevented using this tool for importing more than a few hundred bugs. Fortunately, GitHub provides a (beta) issue import API which, in addition to not triggering notifications, also allows importing one bug, its comments, and meta information such as closed status and assignee in a single request (thus reducing the duration to import 4900 bugs to just a few hours). Furthermore, using this API allows to keep the dates of imported bug reports and comments, which is very useful to this study. We didn't manage to import correct closing dates for every bug report, so we will not study the impact of the bug tracker switch on the time it takes to close a bug.

This improvement has now been integrated upstream as an optional setting.
\end{itemize}

This tool required a correspondence table between Bugzilla and GitHub accounts. Ours was created by manually matching 217 (out of 686) Bugzilla accounts to 175 GitHub accounts,\footnote{The correspondence table can be seen at \url{https://frama.link/vRV9etCd}.} the difference being due to the high prevalence of duplicate Bugzilla accounts. We gave priority to finding GitHub accounts for people who still had opened bug reports.
%The most efficient way to do the matching was by searching the name of the reporter + ``github'' on Google. We also sent e-mails to people we hadn't managed to match (and who still had opened bugs).

Once the switch was approved by the development team on October 4\textsuperscript{th}, 2017, the migration had to happen as soon as possible because every new pull request before the migration added to the number of bug reports that would need to be renumbered. It was conducted %\footnote{More details on the process are available in a blog post at \url{http://theoz.im/bugzilla}. }
on October 18\textsuperscript{th},  the day after the 8.7.0 release (and not before to avoid disturbing the release process). Only 502 out of 4900 bugs (whose numbers were below 1154) had to be renumbered. This number is quite low because these bugs are dating back from the JitterBug era and during the previous switch only the bugs that were still open had been migrated.
Due to the rate of pull request creation, if the switch had been delayed by a year, the number of renumbered bugs would have been closer to 3000.

\section{Data}
\label{data}

@@

All the data for this study was extracted on January 17\textsuperscript{th}, 2019 using the GitHub GraphQL API \cite{github_graphql_API}. Using this API allows to do large requests (100 nodes in a single request) with just the information we need (thus both reducing the bandwidth usage and speeding up the extraction process). We provide the extracted data as CSV files and a Jupyter notebook\footnote{Available at \url{https://github.com/Zimmi48/github-migration-analysis}.} with the code to request this data from GitHub, to load the CSV files, to run the pre-processing steps, and the analyses.%, and the robustness checks.

Migrated bugs and comments have their author information stored in the text, we extract it from there. Dates of bug and comment creation have been preserved which allows us to obtain them transparently for bugs before and after the migration. Bug reports and comments from the JitterBug era are missing information and were not fully migrated, therefore we do not consider any data from before 2008.

@@

\paragraph{Excluding specific reporters}
We have one specific reporter, Jason Gross, who is alone responsible for almost a quarter of all bug reports. To avoid having the behavior of a single individual strongly impact the overall statistics, we exclude his comments, bug reports and the comments they received from our analysis. Similarly, we also exclude the developer who was the main advocate of the bug tracker switch as it could have influenced his behavior before and after the switch.

\paragraph{Merging duplicate Bugzilla accounts}
We found that a significant number of users had created several accounts on the Bugzilla system. We merged them to avoid overestimating the rate of new reporters which is analyzed in the companion Jupyter notebook.

\paragraph{Removal of migration artifact comments}
The migration tool created an issue whose body contains only meta-information, and a first comment with the description, for every bug report. These comments are thus migration artifacts and we remove them from our comment analyses. They can be easily identified because they are the comments that were posted at the exact same date and time as the corresponding bug report.

@@

We measure different indicators of bug tracker activity: the numbers of bug reports per day; the number of distinct bug reporters in a week; the number of new bug reporters (who had never reported a bug before) per day; the number of comments per day; the number of distinct commentators in a week; and the number of new commentators per day. The number of distinct bug reporters and commentators is intended to allow us to distinguish between having a few prolific contributors and a large base of casual contributors. We use an interval of a week instead of a day for measuring the number of distinct bug reporters and commentators because, at the scale of a day, there is less opportunity for repeated contributions by the same contributor, but longer scales would compromise the estimation of effects by removing too much data.

For the visualization of Fig.~\ref{bug_nb_rd} to~\ref{commentator_nb_rd}, we plot a point for the average of each two week period. This is just to allow an easier and clearer visual analysis of the discontinuity. The regression lines, however, correspond to the definitions detailed in the above paragraph (i.e., analysis with daily periods or weekly periods depending on the variable).

We analyze heterogeneous effects by distinguishing between developers and other contributors (others being assimilated to ``users''). We define ``developers'' as the persons who have contributed more than 100 commits since 2008. We identified 18 developers. These developers are responsible for 91.5\% of all commits since 2008 (this is consistent with standard results on the proportion of commits by the ``core team'' in open source software~\cite{robles2009evolution}).

For the dates of the new releases in Section~\ref{descriptive-stats}, we define release dates as the dates of the release announcement on the mailing list or of the corresponding news item on the official website (\url{https://coq.inria.fr/news}), whichever comes first. We selected beta and major releases and excluded release candidates and patch-level releases. The first release to appear in the news section of the current official website is the 8.2 beta release in June, 2008.

\section{Descriptive statistics}
\label{descriptive-stats}

The companion Jupyter notebook contains an extensive array of descriptive tables and figures, such as the distribution of bug reporting in the year and in the day, graphs of bug reporting for individual developers, etc. We only present a subset of these results.

@@

Fig.~\ref{cumulative_bugs} and~\ref{cumulative_comments} give a first (cumulative) view of two of our outcome variables: the number of bug reports, and the number of comments. From these figures, we can clearly notice the dominance of developers in the number of bug comments and the dominance of non-developers in the number of bug reports. We can also already notice some slope changes at the bug tracker switch date. The magnitude and significance of this changes will be assessed in Sec. \ref{results}.

\begin{figure}
\includegraphics{cumulative_bugs.eps}
\caption{Cumulative number of bug reports (since 2008).} \label{cumulative_bugs}
\end{figure}
